---
title: Building a Bodyweight Tracker
author: Vignesh Narayanaswamy
date: '2018-12-31'
slug: building-a-bodyweight-tracker
categories:
  - Shiny
tags:
  - Health Analytics
  - R
  - Shiny
  - RShiny
  - Google Sheets
---
# Overview:
This project is designed to help users track their bodyweight via a google sheet and build an app using Shiny that updates in real time to visualize weightloss progress. Checkout Part 2 (coming soon) of this project which includes a predictive model to help users identify target caloric intake during dieting.

# Steps

1. Create a google sheet
  + Name sheet "Bodyweight Tracker"
2. Create Shiny App


# Shiny App Code

Load the necessary R Packages:
```{r eval=FALSE}
# Define list of required packages
packages<-c("shiny", "googlesheets","DT","ggplot2","ggplot2", "dplyr")
#logical vector that lists which packages are already installed
packages_checked<-(packages %in% installed.packages()[,1])
#list of packages that are not installed
packages_required <- packages[!packages_checked]
#install packages that are required
if (sum(!(packages_checked))>0) {  
  install.packages(packages_required)
}
#load packages
lapply(packages, require, character.only= TRUE)
rm(packages,packages_checked, packages_required)
```

Create Google Sheet object:
```{r eval=FALSE}

bodyweight_sheet <- gs_url("https://docs.google.com/spreadsheets/d/YOURAPIKEY/edit?usp=sharing")

bodyweight_table <- gs_read(bodyweight_sheet)

```

Define Server Logic:
```{r eval=FALSE}

server<- (function(input, output, session) {
  
  #Table of raw data
  output$the_data <- renderDataTable({
    
    datatable(bodyweight_table)
    
  })
  
  
  #Time series plot of bodyweight
  output$the_graph <- renderPlot({
    
    ggplot(data = bodyweight_table, aes(x = as.Date(bodyweight_table$Date, "%m/%d/%Y"), y = bodyweight_table$Bodyweight)) +
      geom_point(color="red")+
      geom_line()+
      stat_smooth(method="auto")+ 
      labs(x = "Date", y="Bodyweight (lbs)")
    
  })
  

    #Time series plot of weekly bodyweight rate of change  
  output$weekly_loss <- renderPlot({
    
    ggplot(data = bodyweight_table, aes(x = as.Date(bodyweight_table$Date, "%m/%d/%Y"), y = bodyweight_table$Weekly_Rate)) +
      geom_point(color="red")+
      geom_line()+
      geom_hline(yintercept=-.02, linetype="dashed", color = "red", size=2)+
      geom_hline(yintercept=-.01, linetype="dashed", color = "red", size=2)+
      stat_smooth(method="auto")+ 
      labs(x = "Date", y="Percent Bodyweight Change Per Week")
    
  })
  
  
  #Tooltip functionality
  output$info <- renderText({
    paste0("x=", input$plot_click$x, "\ny=", input$plot_click$y)
  })
  
})

```

#Define UI Layout:
```{r eval=FALSE}

ui<-(
  fluidPage(
    titlePanel("Bodyweight Tracker"),
    sidebarLayout(
      sidebarPanel(
        tags$a(href="Hyperlink to your Google Sheet", "Click for raw data")
      ),
      mainPanel(
        tabsetPanel(type = "tabs",
                    tabPanel("Daily Bodyweight", plotOutput("the_graph", click = "plot_click"),verbatimTextOutput("info")),
                    tabPanel("Weekly Change", plotOutput("weekly_loss")),
                    tabPanel("Raw Data", dataTableOutput("the_data"))
        )
      )
    )
  ))


# Run the application 
shinyApp(ui = ui, server = server)

```

#Setting up app on Shiny.io

There are 3 main options for using your shiny app

1. Run the app locally (beginner)
2. Run the app remotely by hosting through RStudio's Shiny Server (intermediate)
2. RUn the app remotely by hosting the app on your site (difficult)



<iframe width="1000" height="1000" scrolling="no" frameborder="no"  
src="https://vigneshn.shinyapps.io/bodyweight/"> </iframe>
